---
title: Extending DT child rows example
author: ''
date: '2018-04-13'
slug: extending-dt-child-row-example
categories: []
tags:
  - R
  - DT
---

Short description of R package `DT` from its [website](https://rstudio.github.io/DT/):  

> The R package DT provides an R interface to the JavaScript library [DataTables]((https://datatables.net/)). R data objects (matrices or data frames) can be displayed as tables on HTML pages, and DataTables provides filtering, pagination, sorting, and many other features in the tables.

In practice it means that we can easily publish interactable HTML tables to our reports, dashboards, blog posts etc. To learn more about `DT` visit [https://rstudio.github.io/DT/](https://rstudio.github.io/DT/). As you can imagine, there is lot of stuff you can do with your `datatable`, so even if you have used `datatable()` function for a while, **I'll highly recommend to take a look to the [DT website](https://rstudio.github.io/DT/)**. I found it to be full of useful examples.

# Child Rows

There is [an example](https://rstudio.github.io/DT/002-rowdetails.html) of `datatable` with child rows published in `DT` website. The example, in turn, is adapted from DataTables [website](https://datatables.net/examples/api/row_details.html). In this post I will extend `DT` [child rows example](https://rstudio.github.io/DT/002-rowdetails.html) such that it will look more similar to the original. I start by downloading example data from [DataTables](https://datatables.net/extensions/buttons/examples/initialisation/export.html) website and save it as `dt-export-ex.csv`. Next I will load necessary packages, read data to R and add `Employee ID` variable to data.

```{r fns, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(DT)
x <- readr::read_csv("/home/reigo/blog/dt/DataTables example - File export.csv")
x[["Employee ID"]] <- round(runif(nrow(x)) * 10000)

# datatable2 - datatable with child rows

datatable2 <- function(x, vars = NULL, opts = NULL, ...) {
  
  names_x <- names(x)
  if (is.null(vars)) stop("'vars' must be specified!")
  pos <- match(vars, names_x)
  if (any(map_chr(x[, pos], typeof) == "list"))
    stop("list columns are not supported in datatable2()")
  
  pos <- pos[pos <= ncol(x)] + 1
  rownames(x) <- NULL
  if (nrow(x) > 0) x <- cbind(' ' = '&oplus;', x)

  # options
  opts <- c(
    opts, 
    list(
      columnDefs = list(
        list(visible = FALSE, targets = c(0, pos)),
        list(orderable = FALSE, className = 'details-control', targets = 1),
        list(className = 'dt-left', targets = 1:3),
        list(className = 'dt-right', targets = 4:ncol(x))
      )
  ))
  
  datatable(
    x, 
    ...,
    escape = -2,
    options = opts,
    callback = JS(.callback2(x = x, pos = c(0, pos)))
  )
}

.callback2 <- function(x, pos = NULL) {
  
  part1 <- "table.column(1).nodes().to$().css({cursor: 'pointer'});"
  
  part2 <- .child_row_table2(x, pos = pos)
  
  part3 <- 
  "
   table.on('click', 'td.details-control', function() {
    var td = $(this), row = table.row(td.closest('tr'));
    if (row.child.isShown()) {
      row.child.hide();
      td.html('&oplus;');
    } else {
      row.child(format(row.data())).show();
      td.html('&ominus;');
    }
  });"
  
  paste(part1, part2, part3)
} 

.child_row_table2 <- function(x, pos = NULL) {
  
  names_x <- paste0(names(x), ":")
  text <- "
  var format = function(d) {
    text = '<div><table >' + 
  "

  for (i in seq_along(pos)) {
    text <- paste(text, glue::glue(
        "'<tr>' +
          '<td>' + '{names_x[pos[i]]}' + '</td>' +
          '<td>' + d[{pos[i]}] + '</td>' +
        '</tr>' + " ))
  }

  paste0(text,
    "'</table></div>'
      return text;};"
  )
}

```

```{r, eval=FALSE}
library(tidyverse)
library(DT)
x <- readr::read_csv("dt-export-ex.csv")
x[["Employee ID"]] <- round(runif(nrow(x)) * 10000)
```

I use [`datatable2()`](#datatable2) function (see full code [below](#datatable2)) to embed `datatable` with child rows to this document: 

```{r example-show, message=FALSE, warning=FALSE, eval=FALSE}
datatable2(
  x = x, 
  vars = c("Employee ID", "Age", "Start date"),
  opts = list(pageLength = 5)
)
```

***

```{r example-hide, message=FALSE, warning=FALSE, echo=FALSE}
datatable2(
  x = x, 
  vars = c("Employee ID", "Age", "Start date"),
  opts = list(pageLength = 5)
)
```

When clicking on the + sign we can see additional information about any given row. I think that `datatable` produced with `databale2()` function looks quite similar to [original example](https://datatables.net/examples/api/row_details.html). Also, the benefit of making a function out of it allows us to specify other options available for `DT::datatable()` functions, which hopefully makes `datatable2()` to fit better in my (or yours) workflow.  

# FUNS {#datatable2} 

Code for `datatable2()`:

```{r fns-show, echo=TRUE}
# datatable2 - datatable with child rows
datatable2 <- function(x, vars = NULL, opts = NULL, ...) {
  
  names_x <- names(x)
  if (is.null(vars)) stop("'vars' must be specified!")
  pos <- match(vars, names_x)
  if (any(map_chr(x[, pos], typeof) == "list"))
    stop("list columns are not supported in datatable2()")
  
  pos <- pos[pos <= ncol(x)] + 1
  rownames(x) <- NULL
  if (nrow(x) > 0) x <- cbind(' ' = '&oplus;', x)

  # options
  opts <- c(
    opts, 
    list(
      columnDefs = list(
        list(visible = FALSE, targets = c(0, pos)),
        list(orderable = FALSE, className = 'details-control', targets = 1),
        list(className = 'dt-left', targets = 1:3),
        list(className = 'dt-right', targets = 4:ncol(x))
      )
  ))
  
  datatable(
    x, 
    ...,
    escape = -2,
    options = opts,
    callback = JS(.callback2(x = x, pos = c(0, pos)))
  )
}

.callback2 <- function(x, pos = NULL) {
  
  part1 <- "table.column(1).nodes().to$().css({cursor: 'pointer'});"
  
  part2 <- .child_row_table2(x, pos = pos)
  
  part3 <- 
  "
   table.on('click', 'td.details-control', function() {
    var td = $(this), row = table.row(td.closest('tr'));
    if (row.child.isShown()) {
      row.child.hide();
      td.html('&oplus;');
    } else {
      row.child(format(row.data())).show();
      td.html('&ominus;');
    }
  });"
  
  paste(part1, part2, part3)
} 

.child_row_table2 <- function(x, pos = NULL) {
  
  names_x <- paste0(names(x), ":")
  text <- "
  var format = function(d) {
    text = '<div><table >' + 
  "

  for (i in seq_along(pos)) {
    text <- paste(text, glue::glue(
        "'<tr>' +
          '<td>' + '{names_x[pos[i]]}' + '</td>' +
          '<td>' + d[{pos[i]}] + '</td>' +
        '</tr>' + " ))
  }

  paste0(text,
    "'</table></div>'
      return text;};"
  )
}

```
